### 2014-07-01 Paper1/derivingStiffness.tex

% Title must be 150 characters or less
Define terms: $x$ is hand, $y$ is intent. Capital $F$, $X$ and $Y$ are Laplace transforms. $M$, $B$, and $K$ are mass, damping, and stiffness. $f$ is some externally applied force. Hats ( $\hat{ }$ ) are estimates from an internal model.
\begin{equation}
M\ddot{x}+B\dot{x}+Kx+f=\hat{M}\ddot{y}+\hat{B}\dot{y}+\hat{K}y
\end{equation}
Drop the hats because mass and damping are more or less correctly estimated and we're going to re-represent stiffness misestimation later. Also, switch to the Laplace domain:
\begin{equation}
(Ms^2+Bs+K)X+F=(Ms^2+Bs+K)Y
\end{equation}  
\begin{equation}
F=(Ms^2+Bs+K)(Y-X)
\end{equation}  
Define the quantity, $E=Y-X$, which is essentially error:
\begin{equation}
F=(Ms^2+Bs+K)E
\end{equation} 
Since $M$, $B$, $X$ and $F$ are presumably matters of record, $Y$ and $K$ are the only unknowns. We can extract, but we do so with some as yet unknown stiffness estimation error, $\epsilon$. This means that the stiffness used in the extraction, $K_e=K+\epsilon$.
\begin{equation}
F=(Ms^2+Bs+K+\epsilon)E_e
\end{equation}
So the extracted error, $E_e$ is \textit{not} the same as $E$ because of $\epsilon$. But they're related through $F$:
\begin{equation}
(Ms^2+Bs+K)E=(Ms^2+Bs+K+\epsilon)E_e
\end{equation} 
\begin{equation}
(Ms^2+Bs+K)(E-E_e)=\epsilon E_e
\end{equation} 
\begin{equation}
\frac{E-E_e}{E_e}=\frac{\epsilon}{Ms^2+Bs+K}
\end{equation} 
Generally speaking, this next equation is probably a fair place to stop:
\begin{equation}
\frac{E}{E_e}=1+\frac{\epsilon}{Ms^2+Bs+K}
\end{equation}
But we can re-write it a bit:
\begin{equation}
\frac{Y-X}{Y_e-X}=1+\frac{\epsilon}{Ms^2+Bs+K}
\end{equation} 
\begin{equation}
Y-Y_e=\frac{\epsilon}{Ms^2+Bs+K}(Y_e-X)
\end{equation}
\begin{equation}
\frac{Y-Y_e}{Y_e-X}=\frac{\epsilon}{Ms^2+Bs+K}
\end{equation}
Stiffness misestimation creates a transient in the extraction whose frequency response is entirely defined by the true dynamics of the system. Because we know the value of $K$ used for each extraction, $K_e=\epsilon+K$, we have two unknowns: $Y$ and $\epsilon$. A difference of two different extractions using equation 11 does not contain the $Y$ term:
\begin{equation}
(Y-Y_{1e})-(Y-Y_{2e})=\epsilon_1 \frac{Y_{1e}-X}{Ms^2+Bs+K}- \epsilon_2 \frac{Y_{2e}-X}{Ms^2+Bs+K}
\end{equation}
Because we know the difference in $K_e$ values used for the two extractions ($K_{1e}$ and $K_{2e}$), we know the difference in $\epsilon$ values already.
\begin{equation}
Y_{2e}-Y_{1e}=\epsilon_1 \frac{Y_{1e}-X}{Ms^2+Bs+K}-(\epsilon_1+K_{2e}-K_{1e}) \frac{Y_{2e}-X}{Ms^2+Bs+K}
\end{equation}
\begin{equation}
Y_{2e}-Y_{1e}=\epsilon_1 \frac{Y_{1e}-Y_{2e}}{Ms^2+Bs+K}-(K_{2e}-K_{1e}) \frac{Y_{2e}-X}{Ms^2+Bs+K}
\end{equation}
Because $Y_{1e}$, $Y_{2e}$, and $X$ are known, we can solve for our unknown. In order to deal with the dynamics in the denominators, a slightly lagged value of $K$ should be fine (1 sample = 1 ms delay, $K_{next}=K_{now}-\epsilon_1$). This implies that dynamic tracking of stiffness should be possible in real time. Convergence properties remain unexamined but this might just start and stay at your initial guess (and probably will).
\begin{equation}
(Ms^2+Bs+K)(Y_{2e}-Y_{1e})=\epsilon_1 (Y_{1e}-Y_{2e})-(K_{2e}-K_{1e})(Y_{2e}-X)
\end{equation}
\begin{equation}
(Ms^2+Bs)(Y_{2e}-Y_{1e})=(K+\epsilon_1)(Y_{1e}-Y_{2e})-(K_{2e}-K_{1e})(Y_{2e}-X)
\end{equation}
\begin{equation}
(Ms^2+Bs+K+\epsilon_1)(Y_{2e}-Y_{1e})=(K_{2e}-K_{1e})(X-Y_{2e})
\end{equation}
\begin{equation}
\frac{Y_{1e}-Y_{2e}}{Y_{2e}-X}=\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}
\end{equation}
\begin{equation}
Y_{1e}-Y_{2e}=\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}(Y_{2e}-X)
\end{equation}
Back to zero unknowns, which basically dooms equation 15. Notice that this is equivalent to eq 12.
\begin{equation}
\frac{Y-Y_{2e}}{Y_{2e}-X}-\frac{Y_{1e}-Y_{2e}}{Y_{2e}-X}=\frac{\epsilon_2}{Ms^2+Bs+K}-\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}
\end{equation}
\begin{equation}
\frac{Y-Y_{1e}}{Y_{2e}-X}=\frac{\epsilon_2}{Ms^2+Bs+K}-\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}
\end{equation}

>>>>

% Title must be 150 characters or less
Define terms: $x$ is hand, $y$ is intent. Capital $F$, $X$ and $Y$ are Laplace transforms. $M$, $B$, and $K$ are mass, damping, and stiffness. $f$ is some externally applied force. Hats ( $\hat{ }$ ) are estimates from an internal model.
\begin{equation}
M\ddot{x}+B\dot{x}+Kx+f=\hat{M}\ddot{y}+\hat{B}\dot{y}+\hat{K}y
\end{equation}
Drop the hats because mass and damping are more or less correctly estimated and we're going to re-represent stiffness misestimation later. Also, switch to the Laplace domain:
\begin{equation}
(Ms^2+Bs+K)X+F=(Ms^2+Bs+K)Y
\end{equation}  
\begin{equation}
F=(Ms^2+Bs+K)(Y-X)
\end{equation}  
Define the quantity, $E=Y-X$, which is essentially error:
\begin{equation}
F=(Ms^2+Bs+K)E
\end{equation} 
Since $M$, $B$, $X$ and $F$ are presumably matters of record, $Y$ and $K$ are the only unknowns. We can extract, but we do so with some as yet unknown stiffness estimation error, $\epsilon$. This means that the stiffness used in the extraction, $K_e=K+\epsilon$.
\begin{equation}
F=(Ms^2+Bs+K+\epsilon)E_e
\end{equation}
So the extracted error, $E_e$ is \textit{not} the same as $E$ because of $\epsilon$. But they're related through $F$:
\begin{equation}
(Ms^2+Bs+K)E=(Ms^2+Bs+K+\epsilon)E_e
\end{equation} 
\begin{equation}
(Ms^2+Bs+K)(E-E_e)=\epsilon E_e
\end{equation} 
\begin{equation}
\frac{E-E_e}{E_e}=\frac{\epsilon}{Ms^2+Bs+K}
\end{equation} 
Generally speaking, this next equation is probably a fair place to stop:
\begin{equation}
\frac{E}{E_e}=1+\frac{\epsilon}{Ms^2+Bs+K}
\end{equation}
But we can re-write it a bit:
\begin{equation}
\frac{Y-X}{Y_e-X}=1+\frac{\epsilon}{Ms^2+Bs+K}
\end{equation} 
\begin{equation}
Y-Y_e=\frac{\epsilon}{Ms^2+Bs+K}(Y_e-X)
\end{equation}
\begin{equation}
\frac{Y-Y_e}{Y_e-X}=\frac{\epsilon}{Ms^2+Bs+K}
\end{equation}

Stiffness misestimation creates a transient in the extraction whose frequency response is entirely defined by the true dynamics of the system. Because we know the value of $K$ used for each extraction, $K_e=\epsilon+K$, we have two unknowns: $Y$ and $\epsilon$. A difference of two different extractions using equation 11 does not contain the $Y$ term:
\begin{equation}
(Y-Y_{1e})-(Y-Y_{2e})=\epsilon_1 \frac{Y_{1e}-X}{Ms^2+Bs+K}- \epsilon_2 \frac{Y_{2e}-X}{Ms^2+Bs+K}
\end{equation}
Because we know the difference in $K_e$ values used for the two extractions ($K_{1e}$ and $K_{2e}$), we know the difference in $\epsilon$ values already.
\begin{equation}
Y_{2e}-Y_{1e}=\epsilon_1 \frac{Y_{1e}-X}{Ms^2+Bs+K}-(\epsilon_1+K_{2e}-K_{1e}) \frac{Y_{2e}-X}{Ms^2+Bs+K}
\end{equation}
\begin{equation}
Y_{2e}-Y_{1e}=\epsilon_1 \frac{Y_{1e}-Y_{2e}}{Ms^2+Bs+K}-(K_{2e}-K_{1e}) \frac{Y_{2e}-X}{Ms^2+Bs+K}
\end{equation}
Because $Y_{1e}$, $Y_{2e}$, and $X$ are known, we can solve for our unknown. In order to deal with the dynamics in the denominators, a slightly lagged value of $K$ should be fine (1 sample = 1 ms delay, $K_{next}=K_{now}-\epsilon_1$). This implies that dynamic tracking of stiffness should be possible in real time. Convergence properties remain unexamined but this might just start and stay at your initial guess (and probably will).
\begin{equation}
(Ms^2+Bs+K)(Y_{2e}-Y_{1e})=\epsilon_1 (Y_{1e}-Y_{2e})-(K_{2e}-K_{1e})(Y_{2e}-X)
\end{equation}
\begin{equation}
(Ms^2+Bs)(Y_{2e}-Y_{1e})=(K+\epsilon_1)(Y_{1e}-Y_{2e})-(K_{2e}-K_{1e})(Y_{2e}-X)
\end{equation}
\begin{equation}
(Ms^2+Bs+K+\epsilon_1)(Y_{2e}-Y_{1e})=(K_{2e}-K_{1e})(X-Y_{2e})
\end{equation}
\begin{equation}
\frac{Y_{1e}-Y_{2e}}{Y_{2e}-X}=\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}
\end{equation}
\begin{equation}
Y_{1e}-Y_{2e}=\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}(Y_{2e}-X)
\end{equation}
Back to zero unknowns, which basically dooms equation 15. Notice that this is equivalent to eq 12.
\begin{equation}
\frac{Y-Y_{2e}}{Y_{2e}-X}-\frac{Y_{1e}-Y_{2e}}{Y_{2e}-X}=\frac{\epsilon_2}{Ms^2+Bs+K}-\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}
\end{equation}
\begin{equation}
\frac{Y-Y_{1e}}{Y_{2e}-X}=\frac{\epsilon_2}{Ms^2+Bs+K}-\frac{K_{2e}-K_{1e}}{Ms^2+Bs+K_{1e}}
\end{equation}
